Pure Spin Current Injection in Hydrogenated Graphene Structures

### 2017-05-17 mu/article/article.tex
Changing alpha dependence formulae till line 596

The spin density injection current $\dot{K}^{\mathrm{ab}}$ with speed along
direction $\mathrm{a}$ and spin polarization along $\mathrm{b}$ is defined as
\begin{equation}
\dot{K} = \mu^{\mathrm{abcd}}(\omega)
E^{\mathrm{c}}(\omega) E^{\mathrm{d*}}(\omega)
\label{eq:dotk}
\end{equation}
where
\begin{widetext}
\begin{align}
\mu^{\mathrm{abcd}} (\omega) 
&=
\frac{\pi e^{2}}{\hbar^{2}} \int 
\frac{d^{3}K}{8 \pi^{3}} \sum_{vcc'}^{'}
\mathrm{Re} \left[ K^{\mathrm{ab}}_{cc'} ( 
r^{\mathrm{c}}_{vc'} 
r^{\mathrm{d}}_{cv } +
r^{\mathrm{d}}_{vc'} 
r^{\mathrm{c}}_{cv } ) \right]
\delta(\omega-\omega_{cv})
\label{eq:mu}
\\
K^{\mathrm{ab}}_{mn}
&=
\sum_{\ell}v^{\mathrm{a}}_{nl} S^{\mathrm{b}}_{lm}
\label{eq:velspi-matelem}
\end{align}
\end{widetext}
is the corresponding spin density injection current pseudotensor. The $'$
in the sum means that $c$ and $c'$ are quasi degenerate states and the sum only
covers these states.

Now we define the spin velocity, $\mathcal{V}^{\mathrm{ab}}$ as the speed at
which the spin polarized in the $\mathrm{b}$   direction moves along the
$\mathrm{a}$ direction when a normal incident beam reaches the $xy$ plane with a
polarization angle $\alpha$. Then,
\begin{widetext}
\begin{align}
\mathcal{V}^{\mathrm{ab}} (\omega) 
&= \frac{2}{\hbar}
\frac{\mu^{\mathrm{abxx}}(\omega)
E^{2}(\omega)\cos^{2}(\alpha) + 
\mu^{\mathrm{abyy}}(\omega)
E^{2}(\omega)\sin^{2}(\alpha) + 
2\mu^{\mathrm{abxy}}(\omega)
E^{2}(\omega)\cos(\alpha)\sin(\alpha)}
{\xi^{\mathrm{xx}}(\omega)
E^{2}(\omega)\cos^{2}(\alpha) + 
\xi^{\mathrm{yy}}(\omega)
E^{2}(\omega)\sin^{2}(\alpha)},
\nonumber \\
&= \frac{2}{\hbar}
\frac{\mu^{\mathrm{abxx}}(\omega)\cos^{2}(\alpha) + 
\mu^{\mathrm{abyy}}(\omega)\sin^{2}(\alpha) + 
\mu^{\mathrm{abxy}}(\omega)\sin(2\alpha)}
{\xi^{\mathrm{xx}}(\omega)\cos^{2}(\alpha) + 
\xi^{\mathrm{yy}}(\omega)\sin^{2}(\alpha)}.
\label{eq:vab}
\end{align}
\end{widetext}

@@

\subsection{Fixing spin}\label{sec:theory-fixspin}
Considering that we have 2D structures we can fix the spin direction along the
$x$, $y$, and $z$ Cartesian coordinates and then define the magnitude of the
spin velocity $|\mathcal{V}_{\sigma^{\mathrm{b}}}|$ in a fixed angle
$\gamma_{b}$ as
\begin{align}
|\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega)| 
&=
\sqrt{
[\mathcal{V}^{\mathrm{ax}}(\omega)]^{2}\ +
[\mathcal{V}^{\mathrm{ay}}(\omega)]^{2}\ 
}, 
\label{eq:vs-mag}
\\
\gamma_{\mathrm{b}} (\omega)
&=
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{ay}}(\omega)}
{\mathcal{V}^{\mathrm{ax}}(\omega)} \right),
\label{eq:gamma-ang}
\end{align}
where the angle is measured in the counter-clockwise direction from the positive
$x$ axis.

@@

In a similar way we can fix the velocity in the $xy$ plane
along $x$ and $y$ directions and define $|\mathcal{V}^{\mathrm{a}}|$ as
\begin{equation}
|\mathcal{V}^{\mathrm{a}}(\omega)| = 
\sqrt {
[\mathcal{V}^{\mathrm{ax}}(\omega)]^{2} +
[\mathcal{V}^{\mathrm{ay}}(\omega)]^{2} +
[\mathcal{V}^{\mathrm{az}}(\omega)]^{2} 
},
\label{eq:vv-mag}
\end{equation}
and the corresponding polar and azimuthal angles $\theta$ and $\varphi$ as
\begin{align}
\theta_{\mathrm{a}}  (\omega)
=& 
\cos^{-1} \left( \frac{\mathcal{V}^{\mathrm{az}}(\omega)}
{|\mathcal{V}^{\mathrm{a}}(\omega)|} \right),
& 0 \leq &\theta \leq \pi, 
\label{eq:polar-ang}
\\
\varphi_{\mathrm{a}} (\omega)
=& 
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{ay}}(\omega)}
{\mathcal{V}^{\mathrm{ax}}(\omega)} \right),
& 0 \leq &\varphi \leq 2\pi.
\label{eq:azimuthal-ang} 
\end{align}

@@

\begin{align}
|\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega)|
=& 
\ell_{\mathrm{eff}}
\sum_{\ell=1}^{N_{\mathrm{eff}}}
|\mathcal{V}_{\sigma^{\mathrm{b}}} (\ell | \omega)|
\label{eq:vs-layer}
\\
|\mathcal{V}^{\mathrm{a}}(\omega)|
=&
\ell_{\mathrm{eff}}
\sum_{\ell=1}^{N_{\mathrm{eff}}}
|\mathcal{V}^{\mathrm{a}} (\ell | \omega)|
\label{eq:vv-layer}
\end{align}

@@

We preset the results for $|\mathcal{V}^{\mathrm{a}}(\omega)|$ and
$|\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega)|$ for the C$_{16}$H$_{8}$-alt and
C$_{16}$H$_{8}$-up structures being both noncentrosymmetric semi-infinite 2D
carbon systems with 50\% hydrogenation in different arrangements. The \emph{alt}
system has alternating hydrogen atoms on the upper and bottom sides of the
carbon sheet, while the \emph{up} system has H only on the upper side. We take
the hexagonal carbon lattice to be on the $xy$ plane for both structures, and
the carbon-hydrogen bonds on the perpendicular $xz$ plane, as depicted in Figs.
\ref{fig:alt-struc} and \ref{fig:up-struc}.

We calculated the self-consistent ground state and the Kohn-Sham states using
density functional theory in the local density approximation (DFT-LDA) with a
planewave basis using the ABINIT code \cite{gonzeCPC09}. We used Hartwigsen-
Goedecker-Hutter (HGH) relativistic separable dual-space Gaussian
pseudopotentials \cite{hartwigsenPRB98} including the spin-orbit interaction
needed to calculate $\mu^{\mathrm{abcd}}(\omega)$ presented in Eq.
\eqref{eq:mu}.
% 
It is known that using DFT-LDA to calculate the electronic structure of
materials predict a different band gap than the obtained experimentally. A
correction for the band gap energy value can be calculated by other 
\emph{ab-initio} methods such as the GW approximation \cite{onidaRMP02} but this
is outside the scope of this paper.
% 
The convergence parameters for the calculations of our results corresponding to
the \emph{alt} and \emph{up} structures are cutoff energies of 65\,Ha and
40\,Ha, respectively. The energy eigenvalues and matrix elements were calculated
using 14452 $\mathbf{k}$ points and 8452 $\mathbf{k}$ points in the irreducible
Brillouin zone (IBZ) and present LDA energy band gaps of 0.72\,eV and 0.088\,eV,
respectively for the \emph{alt} and \emph{up} structures.

@@

\begin{table}[b]
\begin{tabular}{cccccc}
\hline
\multirow{2}{*}{Structure \quad} & 
Kind of \quad & 
Pol. &
Energy & 
\multicolumn{2}{c}{$\mathcal{V}^{\mathrm{ab}}(\omega)$}\\
\cline{5-6}
& system & Ang. & [eV] & $\mathrm{ab}$ \quad & [Km/s]\\
\hline
\emph{up}    & 2D   & 40    & 0.09  & $\mathrm{yz}$ &  87.16    \\
             &      &       & 1.94  & $\mathrm{yz}$ &  22.22    \\
             &      &       & 1.97  & $\mathrm{yz}$ & -29.70    \\
\emph{alt}   & 2D   & 145   & 0.72  & $\mathrm{yz}$ & -40.21    \\
             &      &       & 0.91  & $\mathrm{yz}$ & -32.89    \\
 CdSe        & bulk & 90    & 0.91  & $\mathrm{zz}$ & -26.87    \\
 GaAs        & bulk & 90    & 2.31  & $\mathrm{xx}$ & -21.62    \\
\hline
\end{tabular}

@@

In figure \ref{fig:vab-str-comp} we present a comparison of the most intense
responses resulting from evaluate the Eq. \eqref{eq:vab} for the \emph{alt} and
\emph{up} 2D structures vs. CdSe and GaAs bulk systems. As we can see from
this figure the most intense response corresponds to the \emph{up} structure
centered at 0.088\,eV corresponding to the THz radiation and reaching a
spin velocity of 87.2\,Km/s.
% 
In the other hand, for an energy range from 0.66\,eV to 3.0\,eV, corresponding
to energies of the Near Infrared (NIR) and visible radiation, all the four
structures have contributions in the same order of magnitude.
% 
Starting with the 2D structures we have that the \emph{up} structure has other
two peaks centered at 1.94\,eV and 1.97\,eV reaching spin velocities of
22.2\,Km/s and -29.7\,Km/s, respectively, and the \emph{alt} structure has two
peaks centered at 0.72\,eV and 0.91\,eV reaching spin velocities of -40.2\,Km/s
and -32.9\,Km/s, respectively.
% 
Then, for the bulk structures we have that the CdSe has only one intense
response centered at 0.91\,eV reaching a spin velocity of -26.9\,Km/s, and the
GaAs structure has a large and almost planar zone where the response is held
reaching the maximum for an incoming beam of energy of 2.31\,eV and resulting in
a spin velocity of -21.6\,Km/s.
% 
In table \ref{tab:vab-str-comp} we present the comparison of this values for the
2D and bulk structures.
%

@@

    \caption{$|\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega)|$, response
    vs. photon energy vs. polarization angle $\alpha$ for the \emph{up}
    structure for two energy ranges. The absolute maxima is located for an
    energy range from 0.08\,eV to 0.10\,eV and the local maximum from 1.95\,eV
    to 2.0\,eV and both for polarization angles between $25^{\circ}$ and
    $50^{\circ}$.}
    \label{fig:up-3d-vsz}   
\end{figure}

@@

    \caption{Most intense response of
    $|\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega)|$ (top frames, right scale of
    figs (a) and (b)), the corresponding velocity angle
    $\gamma_{\mathrm{z}}(\omega)$ (top frames, right scale), the collinear
    (circled box) and perpendicular (square box) angles, and the two components
    $\mathcal{V}^{\mathrm{xz}}(\omega)$ and $\mathcal{V}^{\mathrm{yz}}(\omega)$
    (bottom frames) for the \emph{up} structure fixing the energy to 0.088\,eV.
    }
    \label{fig:up-vaz-rag}
\end{figure}

% %%%%%%%%%%%%%%%%%
% %% 0.0 eV - 0.16 eV  &  1.80 ev - 2.1 ev
% %% Description of UP |V_{s^z}| 3D
% %%%%%%%%%%%%%%%%%
\textbf{Up structure}

@@

$|\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega)|$ response is obtained and the
second for an energy range from 1.80\,eV to 2.1\,eV, corresponding to visible
radiation, where two local maxima are found.

@@

$|\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega)|$ spectra resulting from evaluate
Eq. \eqref{eq:vs-mag} using different polarization angles $\alpha$ in Eq.
\eqref{eq:vab}. From the figure we have that the onset of the response starts
when the energy of the incoming beam is the same of the gap energy.

@@

$|\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega)|$, related to the left scale, fixing
the energy of the incoming beam to 0.088\,eV and 0.912\,eV, respectively, for
which value the response is maximized for the \emph{up} structure. In the same
figures and frames we present in dashed lines, related to the right scale, the
corresponding velocity angle $\gamma_{\mathrm{z}}(\omega)$ obtained from the
evaluation of Eq. \eqref{eq:gamma-ang}, and in the bottom frames of those
figures the corresponding components $\mathcal{V}^{\mathrm{xz}}(\omega)$ and
$\mathcal{V}^{\mathrm{yz}}(\omega)$. Also we present two circled and square
boxes indicating the values where the angles of the spin velocity are parallel
and perpendicular and the arrows are directed to the value of the response
corresponding to those angles.

@@

$|\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega)|=95.8$\,Km/s coming from the
contribution of the components $\mathcal{V}^{\mathrm{xz}}(\omega)=39.8$\,Km/s
and $\mathcal{V}^{\mathrm{yz}}(\omega)=87.2$\,Km/s for the spin polarized in the
$\mathrm{z}$ direction and having a velocity angle
$\gamma_{\mathrm{z}}(\omega)=65^{\circ}$ on the $xy$ plane.

@@

$\gamma_{\mathrm{z}}(\omega)$ are collinear being both angle $65.5^{\circ}$ and
resulting in a value of the response of
$|\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega)|=82.3$\,Km/s indicated by the upward
green arrow.

@@

$\gamma_{\mathrm{x}}(\omega)=65.5^{\circ}$; for this angles the response has a
value of $|\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega)|=24.8$\,Km/s indicated by
the blue downward arrow.

@@

then, for variations of this magnitude in $\gamma_{\mathrm{z}}(\omega)$ the
complete range of $\alpha$ angle is included because this last is almost

@@

of $|\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega)|=30.3$\,Km/s. This comes from a
major contribution of the $\mathcal{V}^{\mathrm{yz}}(\omega)$ component being
directed in a velocity angle $\gamma_{\mathrm{z}}(\omega)=78^{\circ}$ on the
first Cartesian Quadrant of the $xy$ plane, for the spin polarized in the $z$
direction.

@@

$\alpha$ and the response direction angle $\gamma_{\mathrm{z}}(\omega)$ are
collinear being both $78.5^{\circ}$ and having a response value of
$|\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega)|=23.5$\,Km/s indicated with the
green upward arrow. We found that for variations in the response angle of $\pm
2^{\circ}$, indicated with the error bars, corresponds a range in the

@@

$\gamma_{\mathrm{z}}(\omega)=78.5^{\circ}$ and having a response
$|\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega)|=9.0$\,Km/s indicated with the blue
upward arrow. Again, for variations in the response angle of $\pm2^{\circ}$,
indicated with the error bars, corresponds a range in the polarization angle of
$155^{\circ} \leq \alpha \leq 180^{\circ}$ having again a large range for which
the direction of the response is held.

@@

$|\mathcal{V}_{\sigma^{\mathrm{x}}}(\omega)|=37.4$\,Km/s and
$|\mathcal{V}_{\sigma^{\mathrm{y}}}(\omega)|=24.8$\,Km/s.

@@

>>>>

The spin density injection current $\dot{K}^{\mathrm{ab}}$ with speed along
direction $\mathrm{a}$ and spin polarization along $\mathrm{b}$ is defined as
\begin{equation}
\dot{K}(\omega) = \mu^{\mathrm{abcd}}(\omega)
E^{\mathrm{c}}(\omega) E^{\mathrm{d*}}(\omega)
\label{eq:dotk}
\end{equation}
where
\begin{widetext}
\begin{align}
\mu^{\mathrm{abcd}} (\omega) 
&=
\frac{\pi e^{2}}{\hbar^{2}} \int 
\frac{d^{3}K}{8 \pi^{3}} \sum_{vcc'}^{'}
\mathrm{Re} \left[ K^{\mathrm{ab}}_{cc'} ( 
r^{\mathrm{c}}_{vc'} 
r^{\mathrm{d}}_{cv } +
r^{\mathrm{d}}_{vc'} 
r^{\mathrm{c}}_{cv } ) \right]
\delta(\omega-\omega_{cv})
\label{eq:mu}
\\
K^{\mathrm{ab}}_{mn}
&=
\sum_{\ell}v^{\mathrm{a}}_{nl} S^{\mathrm{b}}_{lm}
\label{eq:velspi-matelem}
\end{align}
\end{widetext}
is the corresponding spin density injection current pseudotensor. The $'$
in the sum means that $c$ and $c'$ are quasi degenerate states and the sum only
covers these states.

Now we define the spin velocity, $\mathcal{V}^{\mathrm{ab}}$ as the speed at
which the spin polarized in the $\mathrm{b}$   direction moves along the
$\mathrm{a}$ direction when a normal incident beam reaches the $xy$ plane with a
polarization angle $\alpha$. Then,
\begin{widetext}
\begin{align}
\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)
&= \frac{2}{\hbar}
\frac{\mu^{\mathrm{abxx}}(\omega)
E^{2}(\omega)\cos^{2}(\alpha) + 
\mu^{\mathrm{abyy}}(\omega)
E^{2}(\omega)\sin^{2}(\alpha) + 
2\mu^{\mathrm{abxy}}(\omega)
E^{2}(\omega)\cos(\alpha)\sin(\alpha)}
{\xi^{\mathrm{xx}}(\omega)
E^{2}(\omega)\cos^{2}(\alpha) + 
\xi^{\mathrm{yy}}(\omega)
E^{2}(\omega)\sin^{2}(\alpha)},
\nonumber \\
&= \frac{2}{\hbar}
\frac{\mu^{\mathrm{abxx}}(\omega)\cos^{2}(\alpha) + 
\mu^{\mathrm{abyy}}(\omega)\sin^{2}(\alpha) + 
\mu^{\mathrm{abxy}}(\omega)\sin(2\alpha)}
{\xi^{\mathrm{xx}}(\omega)\cos^{2}(\alpha) + 
\xi^{\mathrm{yy}}(\omega)\sin^{2}(\alpha)}.
\label{eq:vab}
\end{align}
\end{widetext}

@@

\subsection{Fixing spin}\label{sec:theory-fixspin}
Considering that we have 2D structures we can fix the spin direction along the
$x$, $y$, and $z$ Cartesian coordinates and then define the magnitude of the
spin velocity $|\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)|$ in a fixed angle $\gamma_{\mathrm{b}}(\omega,\alpha)$ as
\begin{align}
|\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)| 
&=
\sqrt{
[\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)]^{2}\ +
[\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)]^{2}\ 
}, 
\label{eq:vs-mag}
\\
\gamma_{\mathrm{b}} (\omega,\alpha)
&=
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)} \right),
\label{eq:gamma-ang}
\end{align}
where the angle is measured in the counter-clockwise direction from the positive
$x$ axis.

@@

In a similar way we can fix the velocity in the $xy$ plane
along $x$ and $y$ directions and define $|\mathcal{V}^{\mathrm{a}}(\omega,\alpha)|$ as
\begin{align}
|\mathcal{V}^{\mathrm{a}}&(\omega,\alpha)| = \nonumber \\
&\sqrt { 
[\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)]^{2} +
[\mathcal{V}^{\mathrm{az}}(\omega,\alpha)]^{2} 
},
\label{eq:vv-mag}
\end{align}
and the corresponding polar and azimuthal angles $\theta$ and $\varphi$ as
\begin{align}
\theta_{\mathrm{a}}  (\omega,\alpha)
=& 
\cos^{-1} \left( \frac{\mathcal{V}^{\mathrm{az}}(\omega,\alpha)}
{|\mathcal{V}^{\mathrm{a}}(\omega,\alpha)|} \right),
& 0 \leq &\theta \leq \pi, 
\label{eq:polar-ang}
\\
\varphi_{\mathrm{a}} (\omega,\alpha)
=& 
\tan^{-1} \left( \frac{\mathcal{V}^{\mathrm{ay}}(\omega,\alpha)}
{\mathcal{V}^{\mathrm{ax}}(\omega,\alpha)} \right),
& 0 \leq &\varphi \leq 2\pi.
\label{eq:azimuthal-ang} 
\end{align}

@@

\begin{align}
|\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)|
=& 
\ell_{\mathrm{eff}}
\sum_{\ell=1}^{N_{\mathrm{eff}}}
|\mathcal{V}_{\sigma^{\mathrm{b}}} (\ell | \omega,\alpha)|
\label{eq:vs-layer}
\\
|\mathcal{V}^{\mathrm{a}}(\omega,\alpha)|
=&
\ell_{\mathrm{eff}}
\sum_{\ell=1}^{N_{\mathrm{eff}}}
|\mathcal{V}^{\mathrm{a}} (\ell | \omega,\alpha)|
\label{eq:vv-layer}
\end{align}

@@

We preset the results for $|\mathcal{V}^{\mathrm{a}}(\omega,\alpha)|$ and
$|\mathcal{V}_{\sigma^{\mathrm{b}}}(\omega,\alpha)|$ for the C$_{16}$H$_{8}$-alt
and C$_{16}$H$_{8}$-up structures being both noncentrosymmetric semi-infinite 2D
carbon systems with 50\% hydrogenation in different arrangements. The \emph{alt}
system has alternating hydrogen atoms on the upper and bottom sides of the
carbon sheet, while the \emph{up} system has H only on the upper side. We take
the hexagonal carbon lattice to be on the $xy$ plane for both structures, and
the carbon-hydrogen bonds on the perpendicular $xz$ plane, as depicted in Figs.
\ref{fig:alt-struc} and \ref{fig:up-struc}.

We calculated the self-consistent ground state and the Kohn-Sham states using
density functional theory in the local density approximation (DFT-LDA) with a
planewave basis using the ABINIT code \cite{gonzeCPC09}. We used 
Hartwigsen-Goedecker-Hutter (HGH) relativistic separable dual-space Gaussian
pseudopotentials \cite{hartwigsenPRB98} including the spin-orbit interaction
needed to calculate $\mu^{\mathrm{abcd}}(\omega,\alpha)$ presented in Eq.
\eqref{eq:mu}.
% 
It is known that using DFT-LDA to calculate the electronic structure of
materials predict a different band gap than the obtained experimentally. A
correction for the band gap energy value can be calculated by other 
\emph{ab-initio} methods such as the GW approximation \cite{onidaRMP02} but this
is outside the scope of this paper.
% 
The convergence parameters for the calculations of our results corresponding to
the \emph{alt} and \emph{up} structures are cutoff energies of 65\,Ha and
40\,Ha, respectively. The energy eigenvalues and matrix elements were calculated
using 14452 $\mathbf{k}$ points and 8452 $\mathbf{k}$ points in the irreducible
Brillouin zone (IBZ) and present LDA energy band gaps of 0.72\,eV and 0.088\,eV,
respectively for the \emph{alt} and \emph{up} structures.

@@

\begin{table}[b]
\begin{tabular}{cccccc}
\hline
\multirow{2}{*}{Structure \quad} & 
Kind of \quad & 
Pol. &
Energy & 
\multicolumn{2}{c}{$\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$}\\
\cline{5-6}
& system & Ang. & [eV] & $\mathrm{ab}$ \quad & [Km/s]\\
\hline
\emph{up}    & 2D   & 40    & 0.09  & $\mathrm{yz}$ &  87.16    \\
             &      &       & 1.94  & $\mathrm{yz}$ &  22.22    \\
             &      &       & 1.97  & $\mathrm{yz}$ & -29.70    \\
\emph{alt}   & 2D   & 145   & 0.72  & $\mathrm{yz}$ & -40.21    \\
             &      &       & 0.91  & $\mathrm{yz}$ & -32.89    \\
 CdSe        & bulk & 90    & 0.91  & $\mathrm{zz}$ & -26.87    \\
 GaAs        & bulk & 90    & 2.31  & $\mathrm{xx}$ & -21.62    \\
\hline
\end{tabular}

@@

{\Large  We calculated the $\mathcal{V}^{\mathrm{ab}}(\omega,\alpha)$} response In figure
\ref{fig:vab-str-comp} we present a comparison of the most intense responses
resulting from evaluate the Eq. \eqref{eq:vab} for the \emph{alt} and \emph{up}
2D structures and CdSe and GaAs bulk systems. As we can see from this figure the
most intense response corresponds to the \emph{up} structure centered at
0.088\,eV corresponding to the Far Infrared (FIR) radiation and reaching a spin
velocity of 87.2\,Km/s.
% 
In the other hand, for an energy range from 0.66\,eV to 3.0\,eV, corresponding
to energies of the Near Infrared (NIR) and visible radiation, all the four
structures have contributions in the same order of magnitude.
% 
Starting with the 2D structures we have that the \emph{up} structure has other
two peaks centered at 1.94\,eV and 1.97\,eV reaching spin velocities of
22.2\,Km/s and -29.7\,Km/s, respectively, and the \emph{alt} structure has two
peaks centered at 0.72\,eV and 0.91\,eV reaching spin velocities of -40.2\,Km/s
and -32.9\,Km/s, respectively.
% 
Then, for the bulk structures we have that the CdSe has only one intense
response centered at 0.91\,eV reaching a spin velocity of -26.9\,Km/s, and the
GaAs structure has a large and almost planar zone where the response is held
reaching the maximum for an incoming beam of energy of 2.31\,eV and resulting in
a spin velocity of -21.6\,Km/s.
% 
In table \ref{tab:vab-str-comp} we present the comparison of this values for the
2D and bulk structures. We found that the most intense response for the spin
velocity of the \emph{up} structure is 3.25 times more intense than CdSe and
4.03 times more intense than the GaAs bulk structures. Also, the \emph{alt}
structure has a response more intense than the bulk systems but being less
intense than the corresponding to the \emph{up} one.
%

@@

    \caption{$|\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha)|$, response
    as a function of the photon energy and polarization angle $\alpha$ for the
    \emph{up} structure for two energy ranges. The absolute maxima is located
    for an energy range from 0.08\,eV to 0.10\,eV and the local maximum from
    1.95\,eV to 2.0\,eV and both for polarization angles between $25^{\circ}$
    and $50^{\circ}$.}
    \label{fig:up-3d-vsz}   
\end{figure}

@@

    \caption{Most intense response of
    $|\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha)|$ (top frames, right
    scale of figs (a) and (b)), the corresponding velocity angle
    $\gamma_{\mathrm{z}}(\omega,\alpha)$ (top frames, right scale), the
    collinear (circled box) and perpendicular (square box) angles, and the two
    components $\mathcal{V}^{\mathrm{xz}}(\omega,\alpha)$ and
    $\mathcal{V}^{\mathrm{yz}}(\omega,\alpha)$ (bottom frames) for the \emph{up}
    structure fixing the energy to 0.088\,eV. }
    \label{fig:up-vaz-rag}
\end{figure}

% %%%%%%%%%%%%%%%%%
% %% 0.0 eV - 0.16 eV  &  1.80 ev - 2.1 ev
% %% Description of UP |V_{s^z}| 3D
% %%%%%%%%%%%%%%%%%

Using the Eq. \eqref{eq:vs-mag}, we calculated the
$|\mathcal{V}_{\sigma^{\mathrm{b}}}|(\omega,\alpha)$ response and made the
analysis for the case when the spin is fixed in the $z$ direction, directed
perpendicularly to the surface of the \emph{alt} and \emph{up} structures. Also,
using the Eq. \eqref{eq:gamma-ang}, we determined the angle
$\gamma_{\mathrm{b}}(\omega,\alpha)$ where the spin-velocity is directed on the
surface of the each structure.

\textbf{Up structure}

@@

$|\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha)|$ response is obtained and
the second for an energy range from 1.80\,eV to 2.1\,eV, corresponding to
visible radiation, where two local maxima are found.

@@

$|\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha)|$ spectra resulting from
evaluate Eq. \eqref{eq:vs-mag} using different polarization angles $\alpha$ in
Eq. \eqref{eq:vab}. From the figure we have that the onset of the response
starts when the energy of the incoming beam is the same of the gap energy.

@@

$|\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha)|$, related to the left scale,
fixing the energy of the incoming beam to 0.088\,eV and 0.912\,eV, respectively,
for which value the response is maximized for the \emph{up} structure. In the
same figures and frames we present in dashed lines, related to the right scale,
the corresponding velocity angle $\gamma_{\mathrm{z}}(\omega,\alpha)$ obtained
from the evaluation of Eq. \eqref{eq:gamma-ang}, and in the bottom frames of
those figures the corresponding components
$\mathcal{V}^{\mathrm{xz}}(\omega,\alpha)$ and
$\mathcal{V}^{\mathrm{yz}}(\omega,\alpha)$. Also we present two circled and
square boxes indicating the values where the angles of the spin velocity are
parallel and perpendicular and the arrows are directed to the value of the
response corresponding to those angles.

@@

$|\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha)|=95.8$\,Km/s coming from the
contribution of the components
$\mathcal{V}^{\mathrm{xz}}(\omega,\alpha)=39.8$\,Km/s and
$\mathcal{V}^{\mathrm{yz}}(\omega,\alpha)=87.2$\,Km/s for the spin polarized in
the $\mathrm{z}$ direction and having a velocity angle
$\gamma_{\mathrm{z}}(\omega,\alpha)=65^{\circ}$ on the $xy$ plane.

@@

$\gamma_{\mathrm{z}}(\omega,\alpha)$ are collinear being both angle
$65.5^{\circ}$ and resulting in a value of the response of
$|\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha)|=82.3$\,Km/s indicated by the
upward green arrow.

@@

$\gamma_{\mathrm{x}}(\omega,\alpha)=65.5^{\circ}$; for this angles the response
has a value of $|\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha)|=24.8$\,Km/s
indicated by the blue downward arrow.

@@

then, for variations of this magnitude in $\gamma_{\mathrm{z}}(\omega,\alpha)$
the complete range of $\alpha$ angle is included because this last is almost

@@

of $|\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha)|=30.3$\,Km/s. This comes
from a major contribution of the $\mathcal{V}^{\mathrm{yz}}(\omega,\alpha)$
component being directed in a velocity angle
$\gamma_{\mathrm{z}}(\omega,\alpha)=78^{\circ}$ on the first Cartesian Quadrant
of the $xy$ plane, for the spin polarized in the $z$ direction.

@@

$\alpha$ and the response direction angle $\gamma_{\mathrm{z}}(\omega,\alpha)$
are collinear being both $78.5^{\circ}$ and having a response value of
$|\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha)|=23.5$\,Km/s indicated with
the green upward arrow. We found that for variations in the response angle of
$\pm 2^{\circ}$, indicated with the error bars, corresponds a range in the

@@

$\gamma_{\mathrm{z}}(\omega,\alpha)=78.5^{\circ}$ and having a response
$|\mathcal{V}_{\sigma^{\mathrm{z}}}(\omega,\alpha)|=9.0$\,Km/s indicated with
the blue upward arrow. Again, for variations in the response angle of
$\pm2^{\circ}$, indicated with the error bars, corresponds a range in the
polarization angle of $155^{\circ} \leq \alpha \leq 180^{\circ}$ having again a
large range for which the direction of the response is held.

@@

$|\mathcal{V}_{\sigma^{\mathrm{x}}}(\omega,\alpha)|=37.4$\,Km/s and
$|\mathcal{V}_{\sigma^{\mathrm{y}}}(\omega,\alpha)|=24.8$\,Km/s.

@@

Now, using the Eq. \eqref{eq:vv-mag}, we calculated the
$|\mathcal{V}^{\mathrm{a}}|(\omega)$ response and made the analysis for the case
when the velocity is fixed in the $x$ and $y$ direction over the surface of the
\emph{alt} and \emph{up} structures. Also, using the Eqns. \eqref{eq:polar-ang}
and \eqref{eq:azimuthal-ang}, we determined the polar
$\theta_{\mathrm{a}}(\omega)$ and azimuthal $\varphi_{\mathrm{a}}(\omega)$
angles where the spin polarization is directed.
